---
title: "geoduck genome linked transcriptome"
author: "Olivia Cattau"
date: '2023-04-06'
output: html_document
---
```{r echo=FALSE}
library(tidyverse)
library(stringr)
library(UpSetR)
library(devtools)
#install_github("jokergoo/ComplexHeatmap")
#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")
#BiocManager::install(version = "3.14")
library(BiocManager)
library(tidyr)
library(UpSetR)
library(ComplexHeatmap)
library(tidyverse)
library(magrittr) # needs to be run every time you start R and want to use %>%
library(dplyr)   
library(scales)
library(unikn)
```

#LOAD DATA
##load new genome linked transcriptome PGEN
```{r}
salmonMatrix<-read.table(file="https://raw.githubusercontent.com/sr320/paper-geoduck-gene-exp/main/data/salmon.merged.gene_tpm.tsv", header=TRUE, sep = '\t')
names(salmonMatrix)[1]<-"target_id"
names(salmonMatrix)[2]<-"gene"
head(salmonMatrix)
#34,947
```

##load geoduck gene annotation file from 
```{r}
mappingMatrix<-read.table(file="https://gannet.fish.washington.edu/Atumefaciens/20220419-pgen-gene_annotation_mapping/20220419-pgen-gene-accessions-gene_id-gene_name-gene_description-alt_gene_description-go_ids.tab", header=TRUE, sep = '\t', fill=TRUE)
colnames(mappingMatrix) <- c("target_id", "identifiers", "gene_id", "gene_name", "annotation", "alt_annotation", "goterms")
#7,736
```

##load goslim mapping from Swiss Prott
[link here](https://github.com/RobertsLab/code/tree/master/r_projects/sam/20230328-pgen-gene_annotation-update) 
```{r}
#double check to make sure this is reading the data correctly
goslims<-read.csv(file="https://raw.githubusercontent.com/RobertsLab/code/master/r_projects/sam/20230328-pgen-gene_annotation-update/outputs/02-goslim-mapping/20230329-pgen-annotations-SwissProt-GO-BP_GOslim.tab", header=TRUE, sep = '\t')
#14,672
```

##download other trasncriptome from clam species
```{r}
#from Mox blasting Manilla/Mercenaria on Pgen
Manilla_Pgen<-read.table(file="https://gannet.fish.washington.edu/gigas/data/p.generosa/Manilla_Pgenenerosa_blastx.tab")
Manila_Pgen_new <- separate(Manilla_Pgen, V2, into = c("gene", "Scaffold"), sep = "::")
names(Manila_Pgen_new)[1]<-"Manila_gene"
names(Manila_Pgen_new)[12]<-"e-value"
# remove unnessisary data 
Manila_Pgen_clean <- select(Manila_Pgen_new, -c(3:11, 13))

#Manila_Pgen_new$accessions <- sub("\\.1$", "", Manila_Pgen_new$accessions)

#761 shared genes

#best Manila Transcriptome
Pgen_Manila<-read.table(file="https://gannet.fish.washington.edu/gigas/data/p.generosa/Pgenerosa_Manilla_db_blastx.tab")
Pgen_Manila_new<- separate(Pgen_Manila, V1, into = c("gene", "Scaffold"), sep = "::") 
names(Pgen_Manila_new)[3]<-"Manila_gene"
names(Pgen_Manila_new)[12]<-"e-value"
# remove unnessisary data 
Pgen_Manila_clean <- select(Pgen_Manila_new, -c(2, 4:11, 13))
#Pgen_Manila_new$accessions <-sub("\\.1$", "", Pgen_Manila_new$accessions)
#657 shared genes

Mercenaria_Pgen<-read.table(file="https://gannet.fish.washington.edu/gigas/data/p.generosa/Mercenaria_Pgenenerosa_blastx.tab")
Mercenaria_Pgen_new<- separate(Mercenaria_Pgen, V2, into = c("gene", "Scaffold"), sep = "::") 
names(Mercenaria_Pgen_new)[1]<-"Mercenaria_gene"
names(Mercenaria_Pgen_new)[12]<-"e-value"
# remove unnessisary data 
Mercenaria_Pgen_clean <- select(Mercenaria_Pgen_new, -c(3:11, 13))
#Mercenaria_Pgen_new$accessions <-sub("\\.1$", "", Mercenaria_Pgen_new$accessions)
#Mercenaria_Pgen_new$accessions <-sub("\\.2$", "", Mercenaria_Pgen_new$accessions)
#6,521 shared genes

#best Mercenaria transcriptome
Pgen_Mercenaria<-read.table(file="https://gannet.fish.washington.edu/gigas/data/p.generosa/Pgenerosa_Mercenaria_db_blastx.tab")
Pgen_Mercenaria_new<- separate(Pgen_Mercenaria, V1, into = c("gene", "Scaffold"), sep = "::") 
names(Pgen_Mercenaria_new)[3]<-"Mercenaria_gene"
names(Pgen_Mercenaria_new)[12]<-"e-value"
Pgen_Mercenaria_clean <- select(Pgen_Mercenaria_new, -c(2, 4:11, 13))
#Pgen_Mercenaria_new$accessions <-sub("\\.1$", "", Pgen_Mercenaria_new$accessions)
#Pgen_Mercenaria_new$accessions <-sub("\\.2$", "", Pgen_Mercenaria_new$accessions)
#5,099 shared 
```

##load genomes from blast output
###needs work
```{r}
Mercenaria_genome<-read.csv(file="https://gannet.fish.washington.edu/gigas/data/p.generosa/Pgenerosa_Mercenaria_db_genome_blastn.tab", sep = '\t', header = FALSE)

Quadrangularis_genome<-read.csv(file="https://gannet.fish.washington.edu/gigas/data/p.generosa/Pgenerosa_quadrangularis_db_genome_blastn.tab", sep = '\t', header = FALSE)

```


#JOIN DATA
##combine all clams for scatterplot by e-value
###needs work
```{r}
#all clam transcriptomes
#Reminder: Merc->Pgen = 6521, Manila->Pgen = 761, Pgen -> Merc = 5,099, Pgen -> Manila = 657
#always join large -> small
compairative_clam_transcripts_right <- left_join(Mercenaria_Pgen_clean, Manila_Pgen_clean, by = "gene") 

compairative_clam_transcripts_left <- left_join(Pgen_Mercenaria_clean, Pgen_Manila_clean, by = "gene") 
```

#join clam transcriptome with goslims
```{r}
Manila_Pgen_goslims <- Manila_Pgen_new %>%
left_join(goslims, by=c("gene"))
Pgen_Manila_goslims <- Pgen_Manila_new %>%
  left_join(goslims, by=c("gene"))
Mercenaria_Pgen_goslims<- Mercenaria_Pgen_new %>%
left_join(goslims, by=c("gene")) #6521
Pgen_Mercenaria_goslims<- Pgen_Mercenaria_new %>%
left_join(goslims, by=c("gene")) #5099
```

#long format
```{r}
tidy_data<-salmonMatrix
head(tidy_data)
tidy_data_long<-tidy_data %>%
  gather(key="tissue", value="tpm", ctenidia:larvae) #name key as time, name value as tpm, gather columns heart:larvae, converts wide to long format
names(tidy_data_long)[2]<-"gene"
head(tidy_data_long)
tidy_data_long$binary=ifelse(tidy_data_long$tpm > 0, 1, 0)

#add goslims below
```

#combine datasets into one gene annotation file***
needs double checks
```{r}
pgenerosa_complete<-salmonMatrix %>% 
  left_join(mappingMatrix, by=c("target_id")) 
# could also do this with goslims
pgenerosa_simple_goslims<-salmonMatrix %>% #best pgenerosa data set
  left_join(goslims, by=c("gene"))
pgenerosa_goslims<-pgenerosa_complete %>%
  left_join(goslims, by=c("gene"))
pgenerosa_long<-tidy_data_long %>% #best long pgenerosa data set
  left_join(goslims, by=c("gene"))
pgenerosa_goslims_long<-tidy_data_long %>%
  left_join(pgenerosa_simple_goslims, by=c("gene"))
```


#break semi-colon in Term column and count unique values only
##distinct goslims
```{r}
# check for duplicate values within each cell using strsplit() and anyDuplicated()
duplicates <- sapply(strsplit(pgenerosa_simple_goslims$gene, ";"), anyDuplicated)

# print the row indices of cells with duplicate values
if (any(duplicates)) {
  print(which(duplicates > 0))
} else {
  print("No duplicates found")
}

#accession numbers = no duplicates
#gene_ID = no duplicates
#allGoID duplicates = column 24458
#BP_GO_ID duplicates = column 24458
#CC_GO_ID duplicates = no duplicates
#MF_GO_ID duplicates = columns 4561  8040 13044 16138
#--------------------------------
#GOslim duplicates lots and lots 
#Terms duplicates lots and lots 

#break cells and reform with unique values only
distinct_goslims <- pgenerosa_simple_goslims %>%
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
  distinct(Term_unique, .keep_all = TRUE) %>%
  mutate(GOslim_unique = lapply(strsplit(as.character(GOslim), ";"),function(GOslim_unique) paste0(unique(GOslim_unique), collapse = ";"))) %>% 
  distinct(GOslim_unique, .keep_all = TRUE)
```

#SPECIES
#Separate the Term columb by semi-colons to count occurances for table
```{r}
# For Pgenerosa alone (wide format) 
# Separate the "Term" column by semi-colons and count the occurrences
term_count_Pgen <- distinct_goslims %>% #make sure you have removed duplicate values beforehand
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) # Count the occurrences and sort by descending count

# Print the count of each term
print(term_count_Pgen)

# For Pgenerosa using pgenerosa_long (versus wide)
# Separate the "Term" column by semi-colons and count the occurrences
term_count_Pgen_long <- pgenerosa_long %>%  # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
  distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) # Count the occurrences and sort by descending count

# Print the count of each term
print(term_count_Pgen_long)
#exactly the same as Pgen_wide


#for Manila Clam
# Separate the "Term" column by semi-colons and count the occurrences
term_count_manila <- Manila_Pgen_goslims %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
  distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) # Count the occurrences and sort by descending count

# Print the count of each term
print(term_count_manila)

# Manila Reverse
term_count_manila_r <- Pgen_Manila_goslims %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
  distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) # Count the occurrences and sort by descending count

# Print the count of each term
print(term_count_manila_r)

#for Mercenaria Clam
# Separate the "Term" column by semi-colons and count the occurrences
term_count_Mercenaria <- Mercenaria_Pgen_goslims %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
  distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) # Count the occurrences and sort by descending count

# Print the count of each term
print(term_count_Mercenaria)

#for Mercenaria Clam reverse
# Separate the "Term" column by semi-colons and count the occurrences
term_count_Mercenaria_r <- Pgen_Mercenaria_goslims %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
  distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) # Count the occurrences and sort by descending count

# Print the count of each term
print(term_count_Mercenaria_r) #over 3x more terms than Mercenaria not reversed...wonder why that is....

#combine all 5 datasets
term_count_all <- bind_rows(term_count_Pgen, term_count_manila, term_count_manila_r, term_count_Mercenaria, term_count_Mercenaria_r, .id = "transcriptome") 

term_count_best <- term_count_all %>% 
  mutate(transcriptome = ifelse(transcriptome == 1, "Pgenerosa", 
                    ifelse(transcriptome == 2, "Manila",
                           ifelse(transcriptome == 3, "Manila_r", 
                                  ifelse(transcriptome == 4, "Mercenaria", 
                                         ifelse(transcriptome == 5, "Mercenaria_r", NA))))))

#write dataframe into a CSV for table making
write.csv(term_count_best, file="/Users/oliviacattau/Documents/GitHub/olivia-geoduck/output/unique_term_count_mult_clam_transcriptomes.csv")
```
##normalize
```{r}
#normalize by Pgen transcriptome
sum_terms<- term_count_Pgen %>% #count terms in n 
  summarise(total_n = sum(n)) #17,611

Pgen_normalized<- term_count_Pgen %>%
  mutate(Pgen_normalized = n/17611)

normal_counts<- term_count_best %>%
  group_by(transcriptome) %>%
  summarise(total_n = sum(n))
#Pgen = 17,611
#Manila = 1235
#Manila_r = 1390
#Mercenaria = 2198
#Mercenaria_r = 7027


filtered_data1 <- term_count_best %>%
  filter(transcriptome %in% c("Mercenaria_r", "Manila", "Mercenaria", "Manila_r")) %>%
  mutate(n_divided = case_when(
                               transcriptome == "Manila" ~ n/1235,
                               transcriptome == "Manila_r" ~ n/1390,
                               transcriptome == "Mercenaria" ~ n/2198,
                               transcriptome == "Mercenaria_r" ~ n/7027))

grouped_data1<-left_join(filtered_data1, Pgen_normalized, by="Term_unique")

species_normal<-grouped_data1 %>%
  mutate(normalized = n_divided/Pgen_normalized)%>%
  rename(n_species = n.x) %>%
  rename(n_Pgen = n.y) %>%
  mutate(log_scale = log(normalized))
#for heatmap

write.csv(species_normal, file="/Users/oliviacattau/Documents/GitHub/olivia-geoduck/output/BEST_species_normalized.csv")
```

##count GOterms unique to species
```{r}
#####make new variables with binary data

#wide format data

Manila_2<-Manila_Pgen_goslims %>%
  select(-c(3:11, 13:23))

Mercenaria_2<-Mercenaria_Pgen_goslims %>%
   select(-c(3:11, 13:23))

species_joined<-left_join(Mercenaria_2, Manila_2, by="gene")

#Mercenaria
species_joined$`e-value.x`=ifelse(species_joined$`e-value.x` > 0, 1, 0)
#Manila
species_joined$`e-value.y`=ifelse(species_joined$`e-value.y` > 0, 1, 0)

species_matrix<-species_joined %>% select(gene, `e-value.x`, `e-value.y`) %>%
  rename(Mercenaria_evalue = `e-value.x`)%>%
  rename(Manila_evalue = `e-value.y`)%>%
  distinct(gene, .keep_all = TRUE)%>%
  mutate(Manila_evalue = ifelse(is.na(Manila_evalue), 0, Manila_evalue))%>%
  mutate(Pgen_evalue = 1)

species_matrix<-tibble::column_to_rownames(species_matrix, var="gene")

species_matrix<-as.matrix(species_matrix) #count matrix for UpSet plots

matrix_species = make_comb_mat(species_matrix)
set_size(matrix_species)
comb_size(matrix_species)

# create sample data frame
species_count<- as.data.frame(species_matrix) %>%
  mutate(Mercenaria_Manila = NA) %>%
  mutate(Manila_Pgen = NA) %>%
  mutate(Pgen_Mercenaria = NA) %>%
  mutate(all=NA)%>%
  mutate(Pgen= NA)

# iterate through rows
########################
#Mercenaria_Manila does not include Pgen genome or else will return all zeros
species_count$Mercenaria_Manila <-ifelse(species_count$Mercenaria_evalue ==1 & species_count$Manila_evalue == 1, 1, 0)

#Manila + Pgenerosa
species_count$Manila_Pgen <-ifelse(species_count$Mercenaria_evalue ==0 & species_count$Manila_evalue == 1 & species_count$Pgen_evalue == 1, 1, 0)

#Mercenaria + Pgenerosa
species_count$Pgen_Mercenaria <-ifelse(species_count$Mercenaria_evalue == 1 & species_count$Manila_evalue == 0 & species_count$Pgen_evalue == 1, 1, 0)

#Mercenaria +Pgenerosa + Manila
species_count$all<-ifelse(species_count$Mercenaria_evalue == 1 & species_count$Manila_evalue == 1 & species_count$Pgen_evalue == 1, 1, 0)

#Pgen alone
species_count$Pgen <-ifelse(species_count$Mercenaria_evalue == 0 & species_count$Manila_evalue == 0 & species_count$Pgen_evalue == 1, 1, 0)

#return column names when species == 1
# identify row indices where Pgen_Mercenaria column equals 1
row_indices_a<- which(species_count$Mercenaria_Manila == 1)
row_indices_b<- which(species_count$Manila_Pgen == 1)
row_indices_c<- which(species_count$Pgen_Mercenaria == 1)
row_indices_d<- which(species_count$all == 1)
row_indices_e<- which(species_count$Pgen == 1)

# extract corresponding row names
genes_Mer_Man <- as.data.frame(rownames(species_count[row_indices_a, ])) 
names(genes_Mer_Man)[1]<-"gene"

genes_Man_Pgen <- as.data.frame(rownames(species_count[row_indices_b, ])) 
names(genes_Man_Pgen)[1]<-"gene"

genes_Mer_Pgen <- as.data.frame(rownames(species_count[row_indices_c, ])) 
names(genes_Mer_Pgen)[1]<-"gene"

genes_all <- as.data.frame(rownames(species_count[row_indices_d, ])) 
names(genes_all)[1]<-"gene"

genes_Pgen <- as.data.frame(rownames(species_count[row_indices_e, ])) 
names(genes_Pgen)[1]<-"gene"

#goslims
goslims_genes_MM<-left_join(genes_Mer_Man, goslims, by="gene")
goslims_gene_ManPgen<-left_join(genes_Man_Pgen, goslims, by ="gene")
goslims_genes_MerPgen<-left_join(genes_Mer_Pgen, goslims, by="gene")
goslims_genes_all<-left_join(genes_all, goslims, by="gene")
goslims_genes_Pgen<-left_join(genes_Pgen, goslims, by="gene")

#Unique GOterms only
#Mercenaria + Manila
MM_count_unique <- goslims_genes_MM %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
 # distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, species= "Mercenaria-Manila")

Man_Pgen_count_unique <- goslims_gene_ManPgen %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
 # distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, species= "Manila-Pgen")

Mer_Pgen_count_unique <- goslims_genes_MerPgen %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
 # distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, species= "Mercenaria-Pgen")

all_count_unique <- goslims_genes_all %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
 # distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, species= "all-3")

Pgen_count_unique <- goslims_genes_Pgen %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
 # distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, species= "Pgen_alone")

species_unique<-bind_rows(Mer_Pgen_count_unique, Man_Pgen_count_unique, MM_count_unique, Pgen_count_unique, all_count_unique)

write.csv(species_unique, file="/Users/oliviacattau/Documents/GitHub/olivia-geoduck/output/BEST_species_unique_GOterms.csv")

```

#TISSUE
#Count Goslims by tissue type
```{r}
#count top GOterms
tissue_goslims <- pgenerosa_long %>%
    group_by(tissue) %>% 
  filter(binary == 1) %>%
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
  distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>%
  separate_rows(Term_unique, sep = ";") %>%
  mutate(Term_unique = trimws(Term_unique)) %>%
  count(Term_unique, tissue, sort = TRUE)

write.csv(tissue_goslims, file="/Users/oliviacattau/Documents/GitHub/olivia-geoduck/output/tissue_counts_unique.csv")
```

#NORMALIZE
##normalize goterm count by % in geoduck transcriptome by % tissue goterm counts by tissue transcriptome and return another % which is the normalized goterm count
```{r}
#add new column to variable term_count_Pgen

#normalize data by % hit in Pgen transcriptome
sum_terms<- term_count_Pgen %>% #count terms in n 
  summarise(total_n = sum(n)) #17,611

Pgen_normalized<- term_count_Pgen %>%
  mutate(Pgen_normalized = n/17611)

#normalize tissue data
normal_counts<- tissue_goslims %>%
  group_by(tissue) %>%
  summarise(total_n = sum(n))
#heart = 16,021
#ctenidia = 15,911
#gonad = 14,715
#juv = 17,277
#larvae = 16,632

filtered_data <- tissue_goslims %>%
  filter(tissue %in% c("ctenidia", "heart", "gonad", "larvae", "juvenile")) %>%
  mutate(n_divided = case_when(tissue == "ctenidia" ~ n/15911,
                               tissue == "heart" ~ n/16021,
                               tissue == "gonad" ~ n/14715,
                               tissue == "larvae" ~ n/16632,
                               tissue == "juvenile" ~ n/17277))

#combine Pgen_normalized with filtered_data and add new column normalized where tissue_n/geoduck_n to produce new normalized value

grouped_data<-left_join(filtered_data, Pgen_normalized, by="Term_unique")

Tis_normal<-grouped_data %>%
  mutate(normalized = n_divided/Pgen_normalized)%>%
  rename(n_tissue = n.x) %>%
  rename(n_Pgen = n.y) %>%
  mutate(log_scale = log(normalized))
#for heatmap

write.csv(Tis_normal, file="/Users/oliviacattau/Documents/GitHub/olivia-geoduck/output/BEST_tissues_normalized.csv")
```


##filter by tissue type
not actually unique goslims, only unique to tissue types
```{r}
tidy1<-pgenerosa_goslims_long %>%
  group_by(tissue) %>%
  filter(tpm > 0) 

ctenidia<-pgenerosa_goslims %>%
  filter(ctenidia > 0 & gonad == 0 & heart == 0 & juvenile == 0 & larvae == 0)
count(ctenidia)
#340 obs (unique to ctenidia)
write.csv(ctenidia, file="/Users/oliviacattau/Documents/GitHub/code-for-Pgenerosa/output/ctenidia_unique.csv")

gonad<-pgenerosa_goslims %>%
  filter(gonad > 0 & ctenidia == 0 & heart == 0 & juvenile == 0 & larvae == 0)
count(gonad)
#119 obs (unique to gonad)
write.csv(gonad, file="/Users/oliviacattau/Documents/GitHub/code-for-Pgenerosa/output/gonad_unique.csv")

heart<-pgenerosa_goslims %>%
  filter(heart > 0 & ctenidia == 0 & gonad == 0 & juvenile == 0 & larvae == 0)
count(heart)
#371 obs (unique to heart)
write.csv(heart, file="/Users/oliviacattau/Documents/GitHub/code-for-Pgenerosa/output/heart_unique.csv")

juvenile<-pgenerosa_goslims %>%
  filter(juvenile > 0 & ctenidia == 0 & heart == 0 & gonad == 0 & larvae == 0)
count(juvenile)
#2,180 obs (unique to juvenile alone)
write.csv(juvenile, file="/Users/oliviacattau/Documents/GitHub/code-for-Pgenerosa/output/juvenile_unique.csv")

larvae<-pgenerosa_goslims %>%
  filter(larvae > 0 & ctenidia == 0 & heart == 0 & gonad == 0 & juvenile == 0)
count(larvae)
#868 obs (unique to larvae alone)
write.csv(larvae, file="/Users/oliviacattau/Documents/GitHub/code-for-Pgenerosa/output/larvae_unique.csv")

juvorlarv<-pgenerosa_goslims %>%
  filter(ctenidia == 0 & gonad == 0 & heart == 0) %>%
  filter(juvenile > 0 & larvae > 0)
#5,022 obs unique to both juveniles or larvae
#1,974 obs unique to juveniles and larvae together

all_unique<-pgenerosa_goslims %>%
   filter(ctenidia > 0 & gonad > 0 & heart > 0 & juvenile > 0 & larvae > 0)
write.csv(all_unique, file="/Users/oliviacattau/Documents/GitHub/code-for-Pgenerosa/output/all_unique.csv")
```

## count GOterms for individual tissues
```{r}
gonad
larvae
heart
ctenidia
juvenile

gonad_count_unique <- gonad %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% # distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, tissue= "gonad")

larvae_count_unique <- larvae %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% # distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, tissue= "larvae")

heart_count_unique <- heart %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% # distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, tissue= "heart")

ctenidia_count_unique <- ctenidia %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% # distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, tissue= "ctenidia")

juvenile_count_unique <- juvenile %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% # distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, tissue= "juvenile")

tissue_all_unique_genes_GOterms<-rbind(heart_count_unique, gonad_count_unique, ctenidia_count_unique, larvae_count_unique, juvenile_count_unique)

write.csv(tissue_all_unique_genes_GOterms, file="/Users/oliviacattau/Documents/GitHub/olivia-geoduck/output/BEST_tissue_unique_GOterms.csv")

```

#organize data for UpSet plot
```{r}
library(dplyr)
#####make new variables with binary data
tidy_data$heart=ifelse(tidy_data$heart > 0, 1, 0)
tidy_data$gonad=ifelse(tidy_data$gonad >0, 1, 0)
tidy_data$ctenidia=ifelse(tidy_data$ctenidia > 0, 1, 0)
tidy_data$larvae=ifelse(tidy_data$larvae > 0, 1, 0)
tidy_data$juvenile=ifelse(tidy_data$juvenile > 0, 1, 0)
new_data<-tidy_data %>% select(target_id,heart, gonad, ctenidia, larvae, juvenile)
new_data<-tibble::column_to_rownames(new_data, var="target_id")
new_data<-as.matrix(new_data) #count matrix for UpSet plots
```

#Make UpSet plot for visualization of overlapping data: 
[link to UpSet R page](https://jokergoo.github.io/ComplexHeatmap-reference/book/upset-plot.html#upset-mode)
```{r}
m1 = make_comb_mat(new_data) #only distinct intersections
m2=make_comb_mat(new_data, mode="intersect")
set_size(m1)
comb_size(m1)
comb_data<-as.data.frame(comb_size(m1))

p1<-UpSet(m1 [comb_degree(m1)<=2], column_title ="All intersections of 2 or less") #intersections 2 or less
p2<-UpSet(m1 [comb_degree(m1)>=3], column_title= "All intersection of 3 or more") #intersections 3 or more
p3<-UpSet(m1, top_annotation = HeatmapAnnotation(
    degree = as.character(comb_degree(m1)),
    "Intersection\nsize" = anno_barplot(comb_size(m1), 
        border = FALSE, 
        gp = gpar(fill = "black"), 
        height = unit(2, "cm")
    ), 
    annotation_name_side = "left", 
    annotation_name_rot = 0), column_title="p. generosa transcriptome distinct intersections") #distinct mode

p5<-UpSet(m2 [comb_degree(m2)<=2], column_title ="All intersections of 2 or less")
p6<-UpSet(m2 [comb_degree(m2)>=3], column_title ="All intersections of 3 or more")
#####testing comb_set sizes
x2<-comb_size(m1)
x3<-as.data.frame(x2)
sum(x3$x2) #34,947 --distinct mode
x4<-comb_size(m2)
x5<-as.data.frame(x4)
sum(x5$x4) #450,489 --intersect mode
```

#Make gene list for unique genes per tissue type
```{r}
# create sample data frame
count_df<- as.data.frame(new_data) %>%
  mutate(heart_gonad = NA) %>%
  mutate(heart_ctenidia = NA) %>%
  mutate(ctenidia_gonad = NA)


# iterate through rows
########################
#heart + gonad
# create a new column for shared data between heart and gonad columns
count_df$heart_gonad <- ifelse(count_df$heart == 1 & count_df$gonad == 1 & count_df$ctenidia == 0 & count_df$juvenile == 0 & count_df$larvae == 0, 1, 0)

#heart + ctenidia
count_df$heart_ctenidia <- ifelse(count_df$heart == 1 & count_df$gonad == 0 & count_df$ctenidia == 1 & count_df$juvenile == 0 & count_df$larvae == 0, 1, 0) 

#ctenidia + gonad
count_df$ctenidia_gonad <- ifelse(count_df$heart == 0 & count_df$gonad == 1 & count_df$ctenidia == 1 & count_df$juvenile == 0 & count_df$larvae == 0, 1, 0) 

#larvae + juvenile
count_df$larv_juv <- ifelse(count_df$heart == 0 & count_df$gonad == 0 & count_df$ctenidia == 0 & count_df$juvenile == 1 & count_df$larvae == 1, 1, 0)

#ctenidia + juvenile 
count_df$ctenidia_juv <- ifelse(count_df$heart == 0 & count_df$gonad == 0 & count_df$ctenidia == 1 & count_df$juvenile == 1 & count_df$larvae == 0, 1, 0)

#heart + juvenile
count_df$heart_juv <- ifelse(count_df$heart == 1 & count_df$gonad == 0 & count_df$ctenidia == 0 & count_df$juvenile == 1 & count_df$larvae == 0, 1, 0)

#gonad + juvenile
count_df$gonad_juv <- ifelse(count_df$heart == 0 & count_df$gonad == 1 & count_df$ctenidia == 0 & count_df$juvenile == 1 & count_df$larvae == 0, 1, 0)

#ctenidia + larvae
count_df$larv_ctenidia <- ifelse(count_df$heart == 0 & count_df$gonad == 0 & count_df$ctenidia == 1 & count_df$juvenile == 0 & count_df$larvae == 1, 1, 0)

#heart + larvae
count_df$larv_heart <- ifelse(count_df$heart == 1 & count_df$gonad == 0 & count_df$ctenidia == 0 & count_df$juvenile == 0 & count_df$larvae == 1, 1, 0)

#gonad + larvae
count_df$larv_gonad <- ifelse(count_df$heart == 0 & count_df$gonad == 1 & count_df$ctenidia == 0 & count_df$juvenile == 0 & count_df$larvae == 1, 1, 0)

#return column names when tissues == 1
# identify row indices where heart_gonad column equals 1
row_indices  <- which(count_df$heart_gonad == 1)
row_indices_2<- which(count_df$heart_ctenidia == 1)
row_indices_3<- which(count_df$ctenidia_gonad == 1)
row_indices_4<- which(count_df$larv_juv == 1)
row_indices_5<- which(count_df$ctenidia_juv == 1)
row_indices_6<- which(count_df$heart_juv == 1)
row_indices_7<- which(count_df$gonad_juv == 1)
row_indices_8<- which(count_df$larv_ctenidia == 1)
row_indices_9<- which(count_df$larv_heart == 1)
row_indices_10<- which(count_df$larv_gonad == 1)


# extract corresponding row names
genes_heart_gonad <- as.data.frame(rownames(count_df[row_indices, ])) 
names(genes_heart_gonad)[1]<-"gene"

genes_heart_ctenidia <- as.data.frame(rownames(count_df[row_indices_2,]))
names(genes_heart_ctenidia)[1]<-"gene"

genes_ctenidia_gonad <- as.data.frame(rownames(count_df[row_indices_3, ])) 
names(genes_ctenidia_gonad)[1]<-"gene"

genes_larv_juv <- as.data.frame(rownames(count_df[row_indices_4, ])) 
names(genes_larv_juv)[1]<-"gene"

genes_ctenidia_juv <-as.data.frame(rownames(count_df[row_indices_5, ])) 
names(genes_ctenidia_juv)[1]<-"gene"

genes_heart_juv <- as.data.frame(rownames(count_df[row_indices_6, ])) 
names(genes_heart_juv)[1]<-"gene"

genes_gonad_juv <- as.data.frame(rownames(count_df[row_indices_7, ])) 
names(genes_gonad_juv)[1]<-"gene"

genes_larv_ctenidia <- as.data.frame(rownames(count_df[row_indices_8, ])) 
names(genes_larv_ctenidia)[1]<-"gene"

genes_larv_heart <- as.data.frame(rownames(count_df[row_indices_9, ])) 
names(genes_larv_heart)[1]<-"gene"

genes_larv_gonad <- as.data.frame(rownames(count_df[row_indices_10, ])) 
names(genes_larv_gonad)[1]<-"gene"

#goslims
goslims_genes_hg<-left_join(genes_heart_gonad, goslims, by="gene")
goslims_gene_hc<-left_join(genes_heart_ctenidia, goslims, by ="gene")
goslims_genes_cg<-left_join(genes_ctenidia_gonad, goslims, by="gene")
goslims_gene_lj<-left_join(genes_larv_juv, goslims, by="gene")
goslims_genes_cj<-left_join(genes_ctenidia_juv, goslims, by="gene")
goslims_genes_hj<-left_join(genes_heart_juv, goslims, by="gene")
goslims_genes_gj<-left_join(genes_gonad_juv, goslims, by="gene")
goslims_genes_cl<-left_join(genes_larv_ctenidia, goslims, by="gene")
goslims_genes_hl<-left_join(genes_larv_heart, goslims, by="gene")
goslims_genes_gl<-left_join(genes_larv_gonad, goslims, by="gene")


#Unique GOterms only
#heart-gonad
hg_count_unique <- goslims_genes_hg %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
 # distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, tissues= "heart-gonad")

#heart-ctenidia
hc_count_unique <- goslims_gene_hc %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
  #distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>%# Count the occurrences and sort by descending count
  mutate(count_unique = n()-1, tissues= "heart-ctenidia")

#ctenidia-gonad
cg_count_unique <- goslims_genes_cg %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
  #distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, tissues= "ctenidia-gonad")

#larvae-juvenile
lj_count_unique <- goslims_gene_lj %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
  #distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, tissues= "larvae-juvenile") #count genes with GOslims and remove 1 (for NAs)

#ctenidia-juvenile
cj_count_unique <- goslims_genes_cj %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
  #distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, tissues= "ctenidia-juvenile") #count genes with GOslims and remove 1 (for NAs)

#heart-juvenile
hj_count_unique <- goslims_genes_hj %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
  #distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, tissues= "heart-juvenile") #count genes with GOslims and remove 1 (for NAs)

#gonad-juvenile
gj_count_unique <- goslims_genes_gj %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
  #distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, tissues= "gonad-juvenile") #count genes with GOslims and remove 1 (for NAs)

#heart + larvae
hl_count_unique <- goslims_genes_hl %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
  #distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, tissues= "heart-larvae") #count genes with GOslims and remove 1 (for NAs)

#ctenidia + larvae
cl_count_unique <- goslims_genes_cl %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
  #distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, tissues= "ctenidia-larvae") #count genes with GOslims and remove 1 (for NAs)

#gonad + larvae
gl_count_unique <- goslims_genes_gl %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
  #distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, tissues= "gonad-larvae") #count genes with GOslims and remove 1 (for NAs)

#term count Juv only
juv_count_unique <- juvenile %>% # remove duplicates from Term column and create Term_unique column
  mutate(Term_unique = lapply(strsplit(as.character(Term), ";"), function(Term_unique) paste0(unique(Term_unique), collapse = ";"))) %>% 
  #distinct(Term_unique, .keep_all = TRUE) %>% 
  filter(!is.na(Term_unique)) %>% # Remove rows with missing (NA) values in the "Term" column
  separate_rows(Term_unique, sep = ";") %>% # Separate the "Term" column by semi-colons
  mutate(Term_unique = trimws(Term_unique)) %>% # Remove any leading/trailing white spaces 
  count(Term_unique, sort = TRUE) %>% # Count the occurrences and sort by descending count 
  mutate(count_unique = n()-1, tissues= "juvenile") #count genes with GOslims and remove 1 (for NAs)

write.csv(juv_count_unique, file="/Users/oliviacattau/Documents/GitHub/olivia-geoduck/data/juv_count_unique.csv")

#combine into one DF
unique_overlapping_genes<-bind_rows(hg_count_unique, hc_count_unique, cg_count_unique, lj_count_unique, cj_count_unique, hj_count_unique, gj_count_unique, cl_count_unique, hl_count_unique, gl_count_unique)

write.csv(unique_overlapping_genes, file="/Users/oliviacattau/Documents/GitHub/olivia-geoduck/output/overlapping_tissues_gene_list_counts.csv")
```

#Make stacked bar-plot for visualizing Unique Terms
```{r}
#filter Terms down a bit 

terms_filtered <- term_count_best %>%
  group_by(Term_unique) %>%
  filter(n() >= 5) %>%
  filter(!is.na("Terms_unique")) %>%
  filter(n >= 10)

#load top 20 hits from Github
top20terms<-read.csv(file="https://raw.githubusercontent.com/course-fish546-2023/olivia-geoduck/main/output/top20terms.csv?token=GHSAT0AAAAAAB64XHVOCLUMSKFNBQEYT6FWZCIKZCA")

compairative_transcriptomes<-top20terms %>%
  filter(transcriptome != "Pgenerosa") %>%
  filter(n >= 15)

stacked_barplot<-ggplot(top20terms, aes(x=transcriptome, y=n, fill=Term_unique))+geom_bar(stat = "identity")+theme_bw()

stacked_barplot2<-ggplot(compairative_transcriptomes, aes(x=transcriptome, y=n, fill=Term_unique))+geom_bar(stat = "identity")+theme_bw()
```

#Heatmap for e-values
```{r}
#Z=e-values, Y=Manilla genes, X=Mercenaria genes

top20_long<-reshape(top20terms, idvar = "Term_unique", timevar = "transcriptome", direction = "wide")

tissues_goterms_unique<-read.csv(file="https://raw.githubusercontent.com/ocattau/code-for-Pgenerosa/main/output/tissue_unique_Pgen_terms_heatmap_data.csv")

tissues2<-tissues_goterms_unique %>%
  filter(tissues != "larvae-juvenile")

#heatmap<-ggplot(pgenerosa_long, aes(x=gene, y=tissue, fill=tpm)) + 
 # geom_tile() #too large

heatmap2<-ggplot(top20terms, aes(x=transcriptome, y=Term_unique, fill=n))+geom_tile()+theme_bw()+ scale_fill_gradient(low="#c2f9ff", high="#003460")

heatmap3<-ggplot(compairative_transcriptomes, aes(x=transcriptome, y=Term_unique, fill=n))+geom_tile()+theme_bw() + scale_fill_gradient(low="#c2f9ff", high="#003460")

heatmap7<- ggplot(tissues2, aes(x=tissues, y=Term_unique, fill=n))+geom_tile()+theme_bw()+ scale_fill_gradient(low="#c2f9ff", high="#003460")+theme(axis.text.x = element_text(angle = 45, hjust = 1) )

Tis_normal_2<-Tis_normal %>%
  filter(Term_unique != "NA")

heatmap8<-ggplot(Tis_normal_2, aes(tissue, y=Term_unique, fill=normalized))+geom_tile()+theme_bw()+theme(axis.text.x = element_text(angle = 45, hjust = 1) )+ scale_fill_gradient2(low = "yellow", mid = "white", high = "red", midpoint = 0.9, na.value = "grey50")#need to set color wheel to white = 1 and red = min and blue = max

p3 <- usecol(c("#E77500", "white", "black"), n = 7)

heatmap9<-ggplot(Tis_normal, aes(tissue, y=Term_unique, fill=log_scale))+geom_tile(color="white")+  scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 0, na.value = "grey50")+theme_bw()+theme(axis.text.x = element_text(angle = 45, hjust = 1) )#need to set color wheel to white = 1 and red = min and blue = max

species_normal_2<-species_normal %>%
  filter(Term_unique != "NA")%>%
  filter(transcriptome == "Manila_r" | transcriptome == "Mercenaria_r")

heatmap10<-ggplot(species_normal_2, aes(transcriptome, y=Term_unique, fill=normalized))+geom_tile(color="white")+  scale_fill_gradient2(low = "red", mid = "white", high = "blue", midpoint = 1, na.value = "grey50")+theme_bw()+theme(axis.text.x = element_text(angle = 45, hjust = 1) )

#barplot?

scatter<-ggplot(distinct_goslims, aes(x=gonad, y=juvenile))+geom_point()+theme_bw()

scatter2<-ggplot(distinct_goslims, aes(x=heart, y=juvenile))+geom_point()+theme_bw()

compairative_clam_transcripts_right

scatter4<-ggplot(compairative_clam_transcripts_right, aes(x=`e-value.x`, y=`e-value.y`))+geom_point()+theme_bw()+xlab("Mercenaria")+ylab("Manila")+labs(title="Pgenerosa blast db")+ theme(plot.title = element_text(hjust = 0.5, margin = margin(b = 10)))#removed 3394 rows containing missing values

scatter5<-ggplot(compairative_clam_transcripts_left, aes(x=`e-value.x`, y=`e-value.y`))+geom_point()+theme_bw()+xlab("Mercenaria")+ylab("Manila")+labs(title="Pgenerosa blast query")+ theme(plot.title = element_text(hjust = 0.5, margin = margin(b = 10)))#Removed 4686 rows containing missing values (`geom_point()`). 
```