---
title: "inspecting output data"
author: "Olivia Cattau"
date: "10/21/2021"
output: html_document
---

```{r}
countMatrix<-read.table(file="https://gannet.fish.washington.edu/gigas/analyses/ctenidia-Trueseq-stranded-mRNA-libraries-C1-NR012_S2_L001_R1_001/abundance.tsv", header=TRUE)
head(countMatrix)
```

```{r}
#do this only once
blast_data<-read.table(file="https://gannet.fish.washington.edu/gigas/Panopea-generosa-uniprot_blastx.tab")
names(blast_data)[1]<-"target_id" #renamed to match countMATRIX data
names(blast_data)[2]<-"identifiers" #renamed to search in swiss prot 
```

```{r}
library(dplyr)
library(tidyr)
#merging target_ids to count Matrix to get identifies for swissprot
generosa_counts<-full_join(blast_data, countMatrix, by = "target_id")
generosa_counts<-generosa_counts[-c(3:12)] # not every target ID has a protein identifier...could be an issue later on....
generosa_counts[complete.cases(generosa_counts),] 
#blast_data<-read.csv(file="/Users/oliviacattau/Documents/analysis/code/generosa_counts.csv")#lost count data due to excel stupidity
generosa_counts$identifiers<-sub("\\.\\d+$", "", generosa_counts$identifiers)
##below code is only to grab identifiers, ignore for now
#cols_to_check = 2
#print (generosa_counts[cols_to_check])
#identifiers_only<-as.data.frame(generosa_counts[2])
#write.csv(identifiers_only, file="/Users/oliviacattau/Documents/analysis/code/pgenerosa/Pgenerosa/NR006_1.csv")
```
# Go to Uniprot website and grab all go terms
```{r}
#loaded into gannet and pulled down, do this only once
uniprot_all_reviewed<-read.csv(file="https://gannet.fish.washington.edu/gigas/uniprot-reviewed_yes.tab", sep='\t', header=TRUE)
```

# How to join tables by identifiers 
```{r}
names(uniprot_all_reviewed)[1]<-"identifiers" #renamed
pgenerosa_proteins<-full_join(generosa_counts, uniprot_all_reviewed, by ="identifiers")
#pgenerosa_cleaned<-pgenerosa_proteins[-c(3:12)] #removed unlabeled columns 
#sort by est_counts, remove na's
pgenerosa_sorted<-pgenerosa_proteins[order(pgenerosa_proteins$est_counts, na.last=TRUE, decreasing=TRUE),]
is.na(pgenerosa_sorted)
pgenerosa_na_removed<-pgenerosa_sorted[complete.cases(pgenerosa_sorted), ]
#make file smaller before exporting later 
```
#clean up data by removing counts < 10
```{r}
pgenerosa_scrubbed<-subset(pgenerosa_na_removed, pgenerosa_na_removed$est_counts >=10)
write.table(pgenerosa_scrubbed, file="/Users/oliviacattau/Documents/GitHub/code/characterize_larval_transciptome/pgenerosa_NR012_1_ctenidia.tab")
```
#push to Gannet
```{r} 
#use command line for this step
rsync -avP /Users/oliviacattau/Documents/GitHub/code/characterize_larval_transciptome/inspecting_NR012_L001_ctenidia.tab ocattau@gannet.fish.washington.edu:/volume2/web/gigas/data/p.generosa/
```


```